Forecasting the 2024 Election: Trump Dominates the South, Harris Leads in the Midwest

### 2024-10-22 paper/paper.qmd
commit text changes

# Additional Data Details
Placeholder

@@

@@

# Visualizing the RMSE and accuracy in a plot with skinnier bars
ggplot(accuracy_rmse_df, aes(x = Metric, y = Value)) +
  geom_bar(stat = "identity", fill = "steelblue", width = 0.2) +  # Adjusted width for skinnier bars
  labs(y = "Value") +
  theme_light()
```
[@fig-barplot-diagnostics] displays the performance of the Bayesian logistic regression model used to predict the outcome of the 2024 presidential election. The model's accuracy is approximately 75%, indicating that 75% of the test set's state-level outcomes were correctly predicted. The RMSE (Root Mean Squared Error) is about 0.5, showing the average error in predicting the winning candidate across states. This suggests the model performs reasonably well, but there is some error in predicting the exact outcome in certain states, which may be due to variations in polling data or other unaccounted factors.

>>>>

# Additional Data Details
Given the nature of the presidential election, the electoral college is more impactful and influential on the possible outcome compared to the nationwide popular votes. Therefore, the majority of the pollsters choose to focus on the battleground states to perform their polling, to gain more insight into the public sentiment specific to certain states. For instance, Sienna/NYT mentions that they have in-state interviewers who call their respondents in states such as Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin. This could explain the discrepancies in our dataset, where there are many missing values from states other than the battleground ones. Furthermore, many pollsters focus on state-wide polls instead of nationwide polls given the nature of the election, resulting in fewer data points or missing data for nationwide poll results. Even though the national popular vote and the results in the electoral college don’t line up a lot of the time, the popular votes can still provide information about issues and opinions that are shaping the election as a whole. Therefore, even if we are focusing on selective states, results from other states and popular votes still remain relevant to outcome prediction.

@@

[@fig-barplot-diagnostics] displays the performance of the Bayesian logistic regression model used to predict the outcome of the 2024 presidential election. The model's accuracy is approximately 75%, indicating that 75% of the test set's state-level outcomes were correctly predicted. The RMSE (Root Mean Squared Error) is about 0.5, showing the average error in predicting the winning candidate across states.

@@

# Visualizing the RMSE and accuracy in a plot with skinnier bars
ggplot(accuracy_rmse_df, aes(x = Metric, y = Value)) +
  geom_bar(stat = "identity", fill = "steelblue", width = 0.2) +  # Adjusted width for skinnier bars
  labs(y = "Value") +
  theme_light()
```
This suggests the model performs reasonably well, but there is some error in predicting the exact outcome in certain states, which may be due to variations in polling data or other unaccounted factors.
